---
title: "Expansion"
author: "Zachary Levine"
date: "08/04/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=FALSE)
```

"Mathematical Models for Assessing Vaccination Scenarios in Several Provinces in Indonesia N. Nuraini, K. Khairudin, P. Hadisoemarto, H. Susanto, A. Hasan, and N. Sumarti"

Expanding old gradient functions from replication assignment to incorporate awarness.

```{r}
library("deSolve")
library("tibble")
theN <- 3565000
make_params <- function(Beta = 0.4215,
                        N = theN,
                        gamma = 0.0876,
                        delta = 0.0028,
                        ##Awarness constant.
                        delta_c = 1/N,
                        ##For each age class
                        deltas = c(0.15, 0.16, 0.627, 1.2, 3.1),
                        v = NULL,
                        q = 0.2,
                        nu = 0.85,
                        ##Sharpness constant
                        k = 3,
                        reinf = 0.05,
                        mu = 0,
                        natural_death = 0
                        ){
  return(c("Beta" = Beta,
              "gamma" = gamma,
              "delta" = delta,
               "N" = N,
              ##N * delta_c = 50,
              "delta_c" = delta_c,
              "q" = q,
              "nu" = nu,
              "k" = k,
           "deltas" = deltas,
              "reinf" = reinf,
              "mu" = mu,
             "natural_death" = natural_death))
}
make_inits <- function(S = theN - I,
                    I = 100,
                    Q = 0,
                    R = 0,
                    D = 0,
                    N = theN){
    return(c(
      "S" = S,
      "I" = I,
      "Q" = Q,
      "R" = R,
      "D" = D))
}
make_beta <- function(betavals = c(0.0251, 0.0355, 0.122, 0.1698, 0.4488)){
  contact_matrix <- matrix(data = 5, nrow = 5, ncol = 5)
  Beta <- matrix(data = NA, nrow = 5, ncol = 5)
  for (i in 1:5){
    for (j in 1:5){
      Beta[[i,j]] <- betavals[[i]]*contact_matrix[i,j]
    }
  }
  return(Beta)
}
make_age_inits <- function(I0= c(5, 45, 35, 25, 35), N = theN){
  ##Cite Dr. Bolker's Feedback on Milestone 1 
  n_age <- length(I0)
 state <- unlist(c(S=(N-I0)/n_age, I=I0, Q=rep(0,n_age), R=rep(0, n_age),
D=rep(0,n_age)))
 return(state)
 
}
paramNames <- c("Beta", "gamma", "delta", "N", "q", "nu", "reinf", "mu", "natural_death")
stateNames <- c("S",  "I", "Q", "R", "D")
run_sim <- function(params = make_params(),
                    use_awareness = FALSE,
                    Beta1 = 0.451,
                    Beta2 = 0.655,
                    Beta3 = 0.5122,
                    Beta4 = 0.71698,
                    Beta5 = 0.7488,
                    betavals = c(Beta1, Beta2, Beta3, Beta4, Beta5),
                    ##A function of time for the vaccination strategy.
                    vaccine_strategy = NULL,
                    use_age_structure = TRUE,
                    age_init = make_age_inits(),
                    init = make_inits(),
                    length = 100){
  ##First decide vaccine strategy.
  if (!is.null(vaccine_strategy)){
      v_func <- vaccine_strategy
  }
  else{
    ##Return the null vaccination strategy.
    v_func <- function(t){
      return(0)
      }
  }
  if (use_age_structure) {
    if (!use_awareness){
      theModel <- function(time, state, parameters){
      S <- as.matrix(state[1:5])   
      I <- as.matrix(state[6:10])    
      Q <- as.matrix(state[11:15])  
      R <- as.matrix(state[16:20])   
      D <- as.matrix(state[21:25])
      with(as.list(c(parameters)), {
        v <- v_func(time)
        v_hat <- nu * v
      theBeta <-  make_beta(betavals = betavals)
      beta_I_sums <- numeric(5)
      for (i in 1:5){
        beta_I_sums[i] <- sum(theBeta[i,]%*%as.matrix(I/N))
      }
      dS <-  natural_death * (N - D) + reinf*R - (beta_I_sums + nu*v + mu)*S
      dI <- beta_I_sums*S - (q + mu) *I
      dQ <- q*I - (gamma + delta + mu) * Q
      dR <- gamma*Q + nu*v*S - (reinf+mu) * R
      dD <- delta * Q
      return(list(c(dS, dI, dQ, dR, dD)))
      })
    }
    }
    else{
     ##Use age structure and awareness. 
      theModel <- function(time, state, parameters){
      S <- as.matrix(state[1:5])   
      I <- as.matrix(state[6:10])    
      Q <- as.matrix(state[11:15])  
      R <- as.matrix(state[16:20])   
      D <- as.matrix(state[21:25])
      with(as.list(c(parameters)), {
        v <- v_func(time)
        v_hat <- nu * v
      theBeta <-  make_beta(betavals = betavals)
      beta_I_sums <- numeric(5)
      ##Do this first
      dD <- delta * Q
    ##Cite Dr. Bolker's Milestone 1 Feedback.
    keep <-  grepl(names(parameters), pattern = paste("deltas",collapse="|"))
    deltas <- parameters[keep]
      for (i in 1:5){
        beta_I_sums[i] <- sum(theBeta[i,]%*%as.matrix(I/(N)))/((1+(dD[i]/deltas[i])^k))
      }
      dS <-  natural_death * (N - D) + reinf*R - (beta_I_sums + nu*v + mu)*S
      dI <- beta_I_sums*S - (q + mu) *I
      dQ <- q*I - (gamma + delta + mu) * Q
      dR <- gamma*Q + nu*v*S - (reinf+mu) * R
      return(list(c(dS, dI, dQ, dR, dD)))
      })
    }
    }
  }
  else{
    if (!use_awareness){
        theModel <- function(time, state, parameters) {
      with(as.list(c(state, parameters)), {
        v <- v_func(time)
        v_hat <- nu * v
        dS <-  natural_death*(N - D) + reinf*R - (Beta*(I)/(N) + nu*v + mu)*S
        dI <- Beta*I*S / N - (q + mu) *I
        dQ <- q *I - (gamma + delta + mu) * Q
        dR <- gamma*Q + nu*v*S - (reinf+mu) * R
        dD <- delta * Q
        return(list(c(dS, dI, dQ, dR, dD)))
      })
    }
    }
    else{
      theModel <- function(time, state, parameters) {
      with(as.list(c(state, parameters)), {
        v <- v_func(time)
        v_hat <- nu * v
        dD <- delta * Q
        dS <-  natural_death*(N - D) + reinf*R - (Beta*(I)/(N*(1+(dD/delta_c)^k)) + nu*v + mu)*S
        dI <- Beta*I*S /(N*((1+(dD/delta_c))^k)) - (q + mu) *I
        dQ <- q *I - (gamma + delta + mu) * Q
        dR <- gamma*Q + nu*v*S - (reinf+mu) * R
        return(list(c(dS, dI, dQ, dR, dD)))
      })
    }
    }
  }
time <- seq(from = 0, to = length, by = 1)
if (use_age_structure){
   res <- as.data.frame(ode(y = age_init,
                       times = time,
                        func = theModel,
                         parms = params))
   
  
}
else{
  res <- as.data.frame(ode(y = init,
                       times = time,
                        func = theModel,
                         parms = params))
  colnames(res) <- c("time", stateNames)
}
return(res)
}

############PLOTTING STARTS HERE
plot_sim <- function(use_age_structure =TRUE,
                     length = 150,
                     use_awareness = FALSE,
                     vaccine_strategy = NULL,
                    Beta1 = 0.0251,
                    Beta2 = 0.0355,
                    Beta3 = 0.122,
                    Beta4 = 0.1698,
                    Beta5 = 0.4488,
                    betavals = c(Beta1, Beta2, Beta3, Beta4, Beta5),
                     params = make_params(),
                     sim = run_sim(use_age_structure = use_age_structure,
                                   length = length,
                                   betavals = betavals,
                                   params = params,
                                   use_awareness = use_awareness,
                                   vaccine_strategy = vaccine_strategy),
                     drop = c("S", "R")){
  library("dplyr")
  library("ggpubr")
  library("tidyr")
  library("directlabels")
  library("lubridate")
  ##Do this first.
  sim <- as_tibble(sim)
  if (!use_age_structure){
    sim$"Immune" <- sim$R
    sim$"Cumulative Immune" <- cumsum(sim$R)
    sim$"Death" <- sim$D
    sim$"Cumulative Deaths" <- cumsum(sim$D)
    ##Drop original columns
    sim <- sim[,!colnames(sim) %in% c("S", "D", "R", "Cumulative Recovered", "Cumulative Deaths")]
  }
  ##This will only work if our data is still in the long form.
  else{
    ##Cite Dr. Bolker's Milestone 1 Feedback.
    throwout <-  grepl(colnames(sim), pattern = paste(drop,collapse="|"))
    sim <- sim[,!throwout]
  }
  ncurves <- length(sim)
  ##If we haven't done this already
  sim <- tidyr::pivot_longer(sim, col = !time)
  sim$"Compartment Type" <- sim$name
  ##Rename columns.
  p <- ggline(data = sim, x= "time", y = "value", color = "Compartment Type", size = 1.5, plot_type = "l") +
    labs(title = "Pandemic simulation",
         x = "Time (days)",
         y = "Count")
  p <- p + scale_x_discrete(breaks = seq(from = 0,
                                         to = length, by = 20),
                            labels = month.name[month(lubridate::today() + days(seq(from = 0, to = length, by = 20)))])
    p
}
compare_vaccine_plot <- function(strat = NULL,
                                 length = 150,
                                 use_age_structure = FALSE,
                                 strat1 = NULL,
                                 strat2 = NULL,
                                 drop = c("S", "R", "D","Q")){
  library("tidyr")
  ##Allow for customs strategies to be coded in for simplicity
  if(!is.null(strat)){
      if (strat == "t_null"){
    strat1 <- function(t){return(0)}
    strat2 <- function(t){return(t)}
      }
    else{
    }
  }
else{
}
      sim1 <- run_sim(use_age_structure = use_age_structure,
                                     length = length,
                                     vaccine_strategy = strat1)
      sim2 <- run_sim(use_age_structure = use_age_structure,
                                     length = length,
                                     vaccine_strategy = strat2)
      if (use_age_structure){
      ##If we're using age structure, do this first before we convert to long form.
      sim1 <- sim1[,!grepl(colnames(sim1), pattern = paste(drop,collapse="|"))]
      sim2 <- sim2[,!grepl(colnames(sim2), pattern = paste(drop,collapse="|"))]
      }
      else{
        sim1 <- sim1[,!colnames(sim1) %in% drop]
        sim2 <- sim2[,!colnames(sim2) %in% drop]
      }
      ##Add the vaccination strategy to the labels of whatever is left.
colnames(sim1) <- c("time", paste0(colnames(sim1)[2:length(colnames(sim1))], ": Vaccination Strategy 1"))
colnames(sim2) <- c("time", paste0(colnames(sim2)[2:length(colnames(sim2))], ": Vaccination Strategy 2"))
    sim1 <- tidyr::pivot_longer(sim1, col = !time)
    sim2 <- tidyr::pivot_longer(sim2, col = !time)
    sim <- dplyr::bind_rows(sim1, sim2)
    sim$"Compartment" <- sim$"name"
    p <- ggline(data = sim, x= "time", y = "value", color = "Compartment", size = 1.5, plot_type = "l") +
    labs(title = "Pandemic simulation",
         x = "Time (days)",
         y = "Count")
  p <- p + scale_x_discrete(breaks = seq(from = 0,
                                         to = length, by = 20),
                            labels = month.name[month(lubridate::today() + days(seq(from = 0, to = length, by = 20)))])
    p
    
}

```


Model output for initial conditions
```{r, warning=FALSE, results='hide',message=FALSE}
plot_sim(use_awareness = TRUE )
q = run_sim(use_awareness = TRUE)
plot(q$I5, pch = 10, type = "o")
title( "Infected Elderly (Age Group Five)")
```



This next few code chunks provide a nice framework for plotting the CN covid data by town, with plenty of options. Meant for checking spacial heterogeniety. From when we were looking at maybe including a spacial componenent, and based on commments in the King et al paper.
```{python}
#This chunk handles python imports
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns
ByTown = pd.read_csv("ConCovidTowns.csv")
ccv = pd.read_csv("ConCOVID.csv")
AgeDeathDat = pd.read_csv("ageDeaths.csv")
```

This is the main code for plotting. 
```{python}
#read in the covid data organized by town.

#These are graph settings for Seaborn. Can ignore if trying to 
#understand code --- they mask
#matplotlib features I'm pretty sure.
rc={'lines.linewidth': 4, 'axes.labelsize': 20, 'axes.titlesize': 18}
sns.set(rc=rc)
sns.set_style("whitegrid", {'axes.grid' : False}) #remove grey grid
sns.set_context("paper") #changes the theme
plt.rcParams['figure.figsize'] = [16, 7] #changes the size


#Function for separating out towns.
def getTown(df: pd.DataFrame, n: int) -> pd.DataFrame:
    '''
    Give the dataframe and a town number.
    Returns all columns for that town.
    '''
    #Rename Date Column 
    df.rename(columns = {'Last update date':'Date'}, inplace = True)
    #Make date a proper datetime object
    df['Date'] = pd.to_datetime(df.Date)
    #Sort by Date
    df.sort_values(by='Date')
    #Reset Index 
    #Grab all elements of big array with this town number
    j = df.loc[df["Town number"] == n]
    #Reset index now that other towns are gone
    j.reset_index(drop=True, inplace=True)
    return(j)
    
def cleandatAge(df: pd.DataFrame,
                    diff = False,
                    loess = False) -> pd.DataFrame:
    ''' Take in a dataframe that sucks and output a 
        dataframe that has the right dates,
        is in the right order,
        is disagregated if diff = True, 
        and is smoothed if loess = True. 
        
        Also fix some column names.
    '''
    df['Date'] = pd.to_datetime(df.Date)
    df = df.sort_values(by='Date')
    j = df.reset_index(drop=True)
    return(j)
    
    
def getPeak(df:pd.DataFrame, n:int, metric:str) -> pd.DataFrame:
    '''
    given original dataframe df, town number n, and metric str, 
    return the day at which the most NEW (metric) were recorded.
    For plotting lines on graphs :D
    '''
    maximum = getTown(ByTown, n)[metric].diff().max()
    q = getTown(ByTown, n).loc[getTown(ByTown, n)[metric].diff() == maximum]
    q.reset_index(drop=True, inplace=True)
    
    #Return the first time the max comes up --- one or two towns reported the
    #same max cases twice, always within a week. So this removes the ambiguity.
    
    return(q.head(1))
 
 
def getFullPeakDF(df:pd.DataFrame, metric:str) -> pd.DataFrame:
    '''
    Given original dataframe df, and a metric (eg Confirmed cases), return a 
    dataframe containing the row in which the most new people were registered 
    as having [metric] eg, testing postitive for Confirmed cases. 
    '''
    MaxDF = getPeak(ByTown, 1, metric)
    for i in range(2,ByTown["Town number"].max()):
        MaxDF = pd.concat([MaxDF, getPeak(ByTown, i, metric)]) 
    
    MaxDF['Date'] = pd.to_datetime(MaxDF.Date)
    MaxDF = MaxDF.sort_values(by='Date')
    MaxDF.reset_index(drop=True, inplace=True)
    
    return(MaxDF)    



def plotTownCases(df: pd.DataFrame,
                  n:int,
                  metric: str,
                  col:str,
                  bins:int = 14,
                  maxcaseline = False,
                  IncludeLegend = True):
    '''
    Give df(big dataframe) and town number (n) and return a plot
    of rolling average over (bins) days for daily change in column (metric)
    '''
    
    UnC = getTown(ByTown, n)[metric].diff()
    
    plt.ylabel("New {}, ({}- day avg)".format(metric, bins))
    plt.xticks(rotation=19)
    if IncludeLegend == True:
        plt.plot(getTown(ByTown, n)["Date"],
         UnC.rolling(bins).mean(),
         label = getTown(ByTown, n)["Town"][0],
         color = col,linewidth= 4)
        plt.legend()
    else:
        plt.plot(getTown(ByTown, n)["Date"],
         UnC.rolling(bins).mean(),
         color = col)
    
        
    if maxcaseline == True:
        plt.axvline(x=getPeak(df,n,metric)._get_value(0,"Date"),
                    ls =':',
                    color= col)
                    
def plotAll(df:pd.DataFrame,
            n:int,
            metric:str,
            col:str = 'b',
            bins:int = 14,
            mcl = True):
    '''
    Plot all the towns and highlight town (n) in (col) with rolling avg over
    9bins) and maybe a line where there are the most new cases (not avgd)
    '''
    for i in range(1, ByTown["Town number"].max()):
        if i == n:
            plotTownCases(ByTown, i, metric, col, maxcaseline = mcl,
                          IncludeLegend = True)
            
        else:
            plotTownCases(ByTown,
                          i,
                          metric,
                          'grey',
                          maxcaseline = False,
                          IncludeLegend = False)
        
#makes raw data into nice data for optimizer :)
def npTownForLM(df:pd.DataFrame, n:int, metric: str) -> np.array:
    #Want np.ndarray with
    q = getTown(df, n)[metric].tail(len(getTown(df, n)[metric])-70)
    q = q.to_numpy()
    return(np.nan_to_num(q))
    
def getAgeGroupDF(df:pd.DataFrame, group: str, bins:int = 14) -> pd.DataFrame:
    '''
    Make the death data nice
    '''
    
    df.rename(columns = {'DateUpdated' : 'Date'}, inplace = True)
    df['Date'] = pd.to_datetime(df.Date)
    df = df.sort_values(by = "Date")
    j = df.loc[df['AgeGroups'] == group] 
    j.reset_index(drop = True, inplace  = True)
    
    dict0 = {"Date": j["Date"][1:], "Deaths": np.nan_to_num(np.diff(j["Total deaths"]))}
    _ =  pd.DataFrame(dict0)
    dict1 = {"Date": _["Date"][bins:], "Deaths": _["Deaths"].rolling(14).mean()[bins:]}

    
    return(pd.DataFrame(dict1))
```

Next, produce a plot to check the above chunks for some of the larger towns:
```{python}
#Plots
plt.figure(figsize=(10,5))

plotTownCases(ByTown, 13, "Confirmed cases", 'g', maxcaseline = True)
plotTownCases(ByTown, 151, "Confirmed cases",'b', maxcaseline = True)
plotTownCases(ByTown, 135, "Confirmed cases", 'y', maxcaseline = True)
plotTownCases(ByTown, 12, "Confirmed cases", 'r', maxcaseline = True)
#plotAll(ByTown, 15, "Confirmed cases")
plotTownCases(ByTown, 83, "Confirmed cases", 'c', maxcaseline = True)
plt.show()
```


```{python}
#dataframe stuff:
agedf = pd.read_csv("agedat.csv")
ageDeaths = pd.read_csv("ageDeaths.csv")
j = cleandatAge(agedf, diff = False)

dict0 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age0_9"])}
df0 = pd.DataFrame(dict0)

dict1 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age10_19"])}
df1 = pd.DataFrame(dict1)

dict2 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age20_29"])}
df2 = pd.DataFrame(dict2)

dict3 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age30_39"])}
df3 = pd.DataFrame(dict3)

dict4 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age40_49"])}
df4 = pd.DataFrame(dict4)

dict5 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age50_59"])}
df5 = pd.DataFrame(dict5)

dict6 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age60_69"])}
df6 = pd.DataFrame(dict6)

dict7 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age70_79"])}
df7 = pd.DataFrame(dict7)

dict8 = {"Date": j["Date"][1:], "Cases":  np.diff(j["cases_age80_Older"])}
df8 = pd.DataFrame(dict8)

plt.plot(df1["Date"],df1["Cases"], label ="10 to 19")
plt.plot(df2["Date"],df2["Cases"], label ="20 to 29")
plt.plot(df3["Date"],df3["Cases"], label ="30 to 39" )
plt.plot(df4["Date"],df4["Cases"], label = "40 to 49")
plt.plot(df5["Date"],df5["Cases"], label ="50 to 59")
plt.plot(df6["Date"],df6["Cases"], label ="60 to 69")
plt.plot(df7["Date"],df7["Cases"], label ="70 to 79")
plt.plot(df8["Date"],df8["Cases"], label = "over 80")
plt.plot()
plt.legend()

agedf.columns
```

```{r}
library(reticulate)
#LOESS SMOOTHING
#Actually really not great. Will compare to rolling average:
#based on code by Ben Bolker on Zulip
dd = py$df0
x = dd$Date
y = dd$Cases
plot(Cases~Date, data=dd)

m = loess(Cases~as.numeric(Date),data=dd)
lines(dd$Date,predict(m))
title("loess smoothing of cases, age 0-9")
```

Compared to rolling avg with two-week bin:

```{python}
#Seaborn titles are very weird in reticulate Rmd. Maybe plot seperatley and 
#include as images for final report.
plt.figure(figsize=(10,5))
plt.plot(df0["Date"], df0["Cases"].rolling(14).mean())
plt.plot(df0["Date"], df0["Cases"], marker = "o", linestyle="None")
plt.title("14-day rolling average of cases, age 0-9")
plt.show()
```

```{python}
#Test code to get data from the age-structured deaths dataset. (auto deagregated)
q = getAgeGroupDF(AgeDeathDat, group = "80 and older", bins = 1)
plt.figure(figsize=(10,5))
plt.plot(q["Date"],q["Deaths"])
```

```{python}
###Make  dataframe for 14-day rolling average dataset of cases
##Have five age groups so cluster:
# 0-9  kids stay together
# 10-19 and 20-29 cluster
# 30-39 and 40-49 cluster
# 50-59 and 60-69 cluster
# 70-79 and 80+ cluster
bins = 14

dict = {"Date": df0["Date"][bins:], "Cases":  df0["Cases"].rolling(14).mean()[bins:]}
df0_9 = pd.DataFrame(dict)

dict = {"Date": df1["Date"][bins:], "Cases":  (df1["Cases"] + df2["Cases"]).rolling(14).mean()[bins:]}
df10_29 = pd.DataFrame(dict)

dict = {"Date": df3["Date"][bins:], "Cases":  (df3["Cases"] + df4["Cases"]).rolling(14).mean()[bins:]}
df30_49  = pd.DataFrame(dict)

dict = {"Date": df1["Date"][bins:], "Cases":  (df5["Cases"] + df6["Cases"]).rolling(14).mean()[bins:]}
df50_69  = pd.DataFrame(dict)

dict = {"Date": df1["Date"][bins:], "Cases":  (df7["Cases"] + df8["Cases"]).rolling(14).mean()[bins:]}
df70_89  = pd.DataFrame(dict)

plt.figure(figsize=(10,5))
plt.plot(df0_9["Date"], df0_9["Cases"], label = "Under 10")
plt.plot(df10_29["Date"], df10_29["Cases"], label = "10-29")
plt.plot(df30_49["Date"], df30_49["Cases"], label = "30-49")
plt.plot(df50_69["Date"], df50_69["Cases"], label = "50-69")
plt.plot(df70_89["Date"], df50_69["Cases"], label = "70+")
plt.legend()
plt.title("Case Count by Age Group (rolling average)")
plt.show()

```


```{python}
#DEATHS 
binsNum = 14

#make dataframes:
df0 = getAgeGroupDF(AgeDeathDat, group = "0-9", bins = binsNum)
df1 = getAgeGroupDF(AgeDeathDat, group = "10-19", bins =binsNum)
df2 = getAgeGroupDF(AgeDeathDat, group = "20-29", bins =binsNum)
df3 = getAgeGroupDF(AgeDeathDat, group = "30-39", bins =binsNum)
df4 = getAgeGroupDF(AgeDeathDat, group = "40-49", bins =binsNum)
df5 = getAgeGroupDF(AgeDeathDat, group = "50-59", bins =binsNum)
df6 = getAgeGroupDF(AgeDeathDat, group = "60-69", bins =binsNum)
df7 = getAgeGroupDF(AgeDeathDat, group = "70-79", bins =binsNum )
df8 = getAgeGroupDF(AgeDeathDat, group = "80 and older", bins = binsNum)


##Case dictionaries
dict = {"Date": df0["Date"], "Deaths":  df0["Deaths"].fillna(0)}
Deaths0_9 = pd.DataFrame(dict)

dict = {"Date": df1["Date"], "Deaths":  (df1["Deaths"] + df2["Deaths"]).fillna(0)}
Deaths10_29 = pd.DataFrame(dict)

dict = {"Date": df3["Date"], "Deaths":  (df3["Deaths"] + df4["Deaths"]).fillna(0)}
Deaths30_49  = pd.DataFrame(dict)

dict = {"Date": df5["Date"], "Deaths":  (df5["Deaths"] + df6["Deaths"]).fillna(0)}
Deaths50_69  = pd.DataFrame(dict)

dict = {"Date": df7["Date"], "Deaths":  (df7["Deaths"] + df8["Deaths"])[:-2].fillna(0)}
Deaths70_89  = pd.DataFrame(dict)

plt.figure(figsize=(10,5))
plt.plot(Deaths0_9["Date"], Deaths0_9["Deaths"], label = "Under 10")
plt.plot(Deaths10_29["Date"], Deaths10_29["Deaths"], label = "10-29")
plt.plot(Deaths30_49["Date"], Deaths30_49["Deaths"], label = "30-49")
plt.plot(Deaths50_69["Date"], Deaths50_69["Deaths"], label = "50-69")
plt.plot(Deaths70_89["Date"], Deaths70_89["Deaths"], label = "70+")
plt.legend()
plt.title("new DEATH Count by Age Group (rolling average)")


Deaths0_9.tail()
Deaths70_89.tail()
plt.show()
```





```{r}
##Testing
library(reticulate)
length(py$Deaths70_89$Deaths)
```

```{r}
#What do the deaths look like for old people in our model before fitting?
#Has awareness and age structure both on.

par = make_params()
par["k"] = 2
par["delta_c"] = 1/(theN**2)

plot(diff(run_sim(c(par, make_beta()),use_awareness = TRUE, length= 100)$D5))
#large part of this is vital dynamics maybe? Turning those off.
par["natural_death_rate"] = 0
plot(diff(run_sim(c(par, make_beta()),use_awareness = TRUE, length= 100)$D5))
#Doesn't change anything.
#cumulative
plot(run_sim(c(par, make_beta()),use_awareness = TRUE)$D5)
```


```{r}
###Make  error function
library(MLmetrics) #Gives nice MLE

errorDeaths =  function(parameters, trace=FALSE, pause=FALSE, plot_it=FALSE){

  #run a simulation for input parameters with output the same length as the dataset
  #subtract one because run_sim is too long by one by default. All dfs same length.
  sim_results = run_sim(params = parameters, length = (length(py$Deaths10_29$Deaths) - 1 ),use_awareness = TRUE)

  if (trace) cat(parameters)
  if (pause) scan(quiet=TRUE)
  if (plot_it)  plot_sim(sim_results)

  s1 = (sim_results$D1 )
  s2 = (sim_results$D2 )
  s3 = (sim_results$D3 )
  s4 = (sim_results$D4 )
  s5 = (sim_results$D5 )

  
   rval <- sqrt( MSE(s1, py$Deaths0_9$Deaths) + 
           MSE(s2, py$Deaths10_29$Deaths) +
           MSE(s3, py$Deaths30_49$Deaths) +
           MSE(s4, py$Deaths50_69$Deaths) +
           MSE(s5, py$Deaths70_89$Deaths))

   if (trace) cat(rval,"\n")
   return(rval) 
    
}

errorCases = function(parameters, plot=FALSE){
  #Back when I was trying to fit to cases I was using this, but
  #fitting to cases is real tough --- I couldn't get a parameter (or combination of parameters) that
  #seemed reasonable to fit diff(Total cases) to. Definitely possible and probably not hard,
  #but vital dynamics making negative diff() in complement of susceptable or similar + people dropping
  #out of infected, cumsum($In) not looking great, and few other things lead me to try with deaths      #instead.
  sim_results = run_sim(params = parameters, length = (length(py$df10_29$Cases) -1),use_awareness = TRUE)

  s1 = (sim_results$I1 + sim_results$Q1 )
  s2 = (sim_results$I2 + sim_results$Q2 )
  s3 = (sim_results$I3 + sim_results$Q3 )
  s4 = (sim_results$I4 + sim_results$Q4 )
  s5 = (sim_results$I5 + sim_results$Q5 )
  
  return(  MSE(s1, py$df0_9$Cases) + 
           MSE(s2, py$df10_29$Cases) +
           MSE(s3, py$df30_49$Cases) +
           MSE(s4, py$df50_69$Cases) +
           MSE(s5, py$df70_89$Cases) )
}
```

```{r plot_check}
## BMB machinery to plot sim results and data together to see what the heck is going on
age_brks <- c(0,10,30,50,70,90)
age_cats <- sprintf("%d_%d", head(age_brks,-1), age_brks[-1]-1)
names(age_cats) <- age_cats ## for purrr .id
tmpf <- function(type="Cases") {
    dname <- switch(type, Cases="df", Deaths="Deaths")
    x <- purrr:::map_dfr(age_cats, ~py[[paste0(dname,.)]], .id="agecat")
    names(x)[names(x)==type] <- "value"
    x[["type"]] <- type
    x <- (x
        %>% group_by(type,agecat)
        %>% arrange(Date)
        %>% mutate(time=seq(n()))
        %>% select(-Date)
    )
    return(x)
}
get_data <- function() {
    purrr::map_dfr(c("Cases","Deaths"), tmpf)
}
check_deaths <- . %>% filter(type=="Deaths") %>% arrange(agecat)
get_sim_data <- function(x, diff_deaths=TRUE) {
    dd <- (x
        %>% pivot_longer(-time,names_to="state")
        %>% separate(state,into=c("state","age"),sep=1, convert=TRUE)
        %>% filter(state %in% c("I","Q","D"))
        %>% mutate(agecat=age_cats[age],
                   type=ifelse(state %in% c("I","Q"), "Cases", "Deaths"))
        %>% group_by(time,type,agecat)
        %>% summarise(across(value,sum), .groups="drop")  ## this adds I and Q together
    )
    if (diff_deaths) {
        ## convert long-to-wide, modify Deaths, convert wide-to-long
        dd2 <- (dd
            %>% pivot_wider(names_from=type,values_from=value)
            %>% arrange(agecat)
            %>% group_by(agecat)
            ## %>% mutate(across(Deaths, ~c(0,diff(.))))
            %>% mutate(Deaths=c(0,diff(Deaths)))
            %>% ungroup()
        )
        dd <- dd2  %>% pivot_longer(-c(time,agecat), names_to="type")
    }
    return(dd)
}
g <- get_sim_data(sim_results)

library(ggplot2); theme_set(theme_bw())
pydat <-  get_data()

#' plot simulations and data by age class and deaths/cases
#' @param simres simulation results
#' @param params optional parameters: if specified, will run simulation first
#' @param diff_deaths difference death variable from sims?
#' @param ... extra arguments to pass to \code{run_sim}
plot_simdata <- function(simres, diff_deaths=TRUE, params=NULL, ...) {
    if (!is.null(params)) {
        simres <- run_sim(params = params, length = (length(py$df10_29$Cases) -1), ...)
    }
    
    alldat <- bind_rows(list(data=get_data(),
                             sim=get_sim_data(simres, diff_deaths=diff_deaths)),
                        .id="source")
    ggplot(alldat, aes(time,value,linetype=source, colour=source)) +
        geom_line() +
        facet_grid(type~agecat,scale="free") +
        scale_y_log10()
}
```

```{r}
###Test these out

#Make parameter list
Para1 = make_params()

#Change value for beta
Para1['Beta'] = 0.42

#Change value for gamma
Para1['gamma'] = 0.0082

#Print list with new values as check
print(Para1)

plot_simdata(params=Para1, use_awareness=TRUE)

#Calculate error functions at these new values
errorCases(Para1)
errorDeaths(Para1)

#Run a simulation with these parameters to check for weird infinities like optim claims
sim_results = run_sim(params = Para1, length = (length(py$df10_29$Cases) -1), use_awareness=TRUE)

plot_simdata(sim_results)
#Just use default length --- maybe the issue is with length
#run_sim(params = Para1)
#nope

#check results --- these never change for me. big issue.
#print(sim_results)

#check that the issue isn't with the way I was summing.
#print((sim_results$I1 + sim_results$Q1 + py$df10_29$Cases))
```


```{r}
##Using Optim, fit the model
library(stats)

#make new param list with default params
initparams = make_params()

#change any value to make sure this works; optim should change anyways.
initparams['gamma'] = 0.2

#test that I don't get an infinite value error with these params without optim
errorDeaths(c(make_params(), make_beta()))

#run optim
results = optim(par =  c(make_params(), make_beta() ),fn = errorDeaths, method = "Nelder-Mead", control = list(maxit = 20))
#, upper = c(1,1,1,1,50/40000000,1,1,8,50,50,50,50,50,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1))

#look at results
results
```

```{r}
#Check to see what parameters it changed
results$par == c(make_params(),make_beta())

#Look at parameters
para = results$par
results
```

```{r}
#Run simulation with new parameters
par['k'] = 2
sim_results = run_sim(params = par,  use_awareness = TRUE,length = (length(py$df10_29$Cases)))
plot(diff(run_sim(c(para, make_beta()),use_awareness = TRUE, length= 100)$D5))

para == par
```

```{python}
#Just a bunch of quick plots to look at results
#plot(NewCases2)
res = r.sim_results 

plt.plot(np.diff(res["D1"]))
plt.plot(np.diff(res["D2"]))
plt.plot(np.diff(res["D3"]))
plt.plot(np.diff(res["D4"]))
plt.show()
```

```{python}
plt.plot((res["S5"]))
```

```{python}
plt.plot(Deaths50_69["Deaths"], marker = 'o', linestyle = 'none')
plt.plot(np.diff(res["D4"]))
#plot(sim_results$I1)
```





```{r}
#testing?
#run_sim(params = initparams)
#run_sim(make_params())
```



